Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems

### 2021-02-04 paper/main.tex
5.3

\subsection{Analysis of UMM and CEGO on LOP and PFSP}
To  understand the behavior of the algorithms, we record the fitness of
the solution evaluated at each step of each run and we average those values
over 10 runs with different random seeds on the same problem instance.  As an
example, each line in Fig.~\ref{fig:results} shows the mean fitness of the
solutions evaluated by either UMM or CEGO on one instance. The
shaded area shows one standard deviation around the mean. Ideally, each new
evaluation will monotonically improve in fitness, since each solution evaluated
helps to refine the model and leads to a better solution being evaluated in the
next step.  However, it is possible that new solutions are worse
than their predecessors. %Nevertheless, in this example, there is a clear
% overall improvement as more evaluations are performed.
The minimum values
achieved in the plot up to a particular number of function evaluations gives an
estimation of the mean fitness of the best-so-far solution up to that point in
the run. The algorithms do not return the final solution shown but rather the
best one (minimum fitness) of each run. This result summarization is not the classical. However, due to the limited number of function evaluations, we can have a better picture of the behaviour of the algorithm at each step. As a consequence, results displayed in the figures are not comparable with the original paper of CEGO: for this comparison we provide in Table~\ref{table:fitness} the summary of the best solutions found by each algorithm, which are similar to the original study.

@@

\begin{figure*}[tb]
\small
\centering%
\begin{minipage}{0.49\linewidth}
  \texttt{N-t59b11xx} (LOP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_lop_IO_N-t59b11xx}
\end{minipage}
\begin{minipage}{0.49\linewidth}
  \texttt{N-sgb75-01} (LOP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_lop_SGB_N-sgb75_02}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
  \texttt{N-p40-01} (LOP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_lop_RandB_N-p40-01}
\end{minipage}
\begin{minipage}{0.49\linewidth}
  \texttt{N-p50-01} (LOP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_lop_RandB_N-p50-01}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
  \texttt{reC19} (PFSP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_pfsp_rec19_txt}
\end{minipage}
\begin{minipage}{0.49\linewidth}
  \texttt{reC31} (PFSP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_pfsp_rec31_txt}
\end{minipage}
\label{fig:results}
\caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on various instances.\label{fig:rec05}}
\end{figure*}

>>>>

\subsection{Experimental Analysis of UMM and CEGO} % on LOP and PFSP instances}

To better analyze the behavior of the algorithms, we record the fitness of the
solution evaluated at each step of each run. For each problem instance and each
algorithm, we plot the mean fitness at each evaluation over 10 runs with
different random seeds, as shown in Fig.~\ref{fig:results}.
%
The shaded area shows one standard deviation around the mean. Ideally, each new
evaluation will monotonically improve in fitness, since each solution evaluated
helps to refine the model and leads to a better solution being evaluated in the
next step.  However, it is possible that new solutions are worse than their
predecessors. %Nevertheless, in this example, there is a clear
% overall improvement as more evaluations are performed.

We chose to plot the fitness of each evaluated solution rather than the fitness
of the best-so-far to better understand the behavior of the algorithms at each
step. Nevertheless, the minimum values achieved in the plot up to a particular
number of function evaluations gives an estimation of the mean fitness of the
best-so-far solution up to that point in the run. The algorithms do not return
the final solution shown but rather the best one (minimum fitness) of each run.

This result summarization is not the classical. However, due to the limited number of function evaluations, we can have a better picture of the behaviour of the algorithm at each step. As a consequence, results displayed in the figures are not comparable with the original paper of CEGO: for this comparison we provide in Table~\ref{table:fitness} the summary of the best solutions found by each algorithm, which are similar to the original study.

@@

\begin{figure*}[tb]
\small
\centering%
\begin{minipage}{0.49\linewidth}
  \texttt{N-t59b11xx} (LOP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_lop_IO_N-t59b11xx}
\end{minipage}
\begin{minipage}{0.49\linewidth}
  \texttt{N-sgb75-01} (LOP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_lop_SGB_N-sgb75_02}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
  \texttt{N-p40-01} (LOP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_lop_RandB_N-p40-01}
\end{minipage}
\begin{minipage}{0.49\linewidth}
  \texttt{N-p50-01} (LOP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_lop_RandB_N-p50-01}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
  \texttt{reC19} (PFSP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_pfsp_rec19_txt}
\end{minipage}
\begin{minipage}{0.49\linewidth}
  \texttt{reC31} (PFSP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_pfsp_rec31_txt}
\end{minipage}
\label{fig:results}
\caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on various instances.\label{fig:results}}
\end{figure*}
